---
title: "SLF - Home Care methodology"
author: "James McMahon"
date: "11/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Methodology

### Data import

  1. Extract all data directly from the DVRPOD / social_care_2 database
  1. Fix 2017 issues
    * Fix `period` = "2017" to `period` = "2017Q4"
    * Fix `financial_quarter` = 0 to `financial_quarter` = 4 (for 2017 only)
  1. Set any missing `hc_service` to 0 (Not submitted)
  1. Drop any records flagged with `hc_start_date_after_end_date`
  1. Don't import any unvalidated records, currently up to `r latest_validated_period`
 
### Data cleanup

  1. Where start date is missing replace this with the start of the quarter
  1. Where a CHI has multiple `social_care_id`s for the same `sending_location` replace all ids with the latest submitted one. 
  1. Fill in `reablement` - First group across; `chi, sending_location_name, sending_location, social_care_id, hc_service_start_date, hc_service` which will ultimately be the episode.
    * Where `reablement` = 9 or missing, fill in from other records in the 'episode' 
    * first fill in from the latest records, to the earliest
    * Then from earlier records if still missing.
  1. Replace `hc_service_provider`, when it changes within the episode, to 5 (other).

### Merge data

  1. Merge records together, by the grouping mentioned above, records are the same epsiode if they are for the same person, same service type and start on the same date.
    1. If `reablement` changes within the episode, split the episode at the quarter date between the change. ** We could replicate this logic for `hc_service_provider` **
    1. For Argyll and Bute 2020/21 use `multistaff_input * total_staff_home_care_hours` where `multistaff_input` must be at least 1, for the hours.
    1. For Angus 2018Q3-2019Q3 inclusive recalculate the hours as they were submitted as average weekly hours, so divide by 7 (days) then multiply by the number of days in the quarter to get quarterly hours.
    1. Take the hours for each quarter and create a new variable which is a sum of all hours for that episode in that quarter
  1. Where an episode is split (due to reablement changing) recalculate the dates so that they start/end at the start and end of the quarter where the 'change' happens.

```{r sc_id, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
replaced_start_dates %>% group_by(sending_location_name) %>% summarise(before = n_distinct(social_care_id)) %>% left_join(fixed_sc_ids %>% group_by(sending_location_name) %>% summarise(after = n_distinct(social_care_id))) %>% mutate(diff = before - after, diff_pct = scales::percent(diff/ before, accuracy = 0.1)) %>% gt::gt()
```
